Two NIRCam channels are Better than One: How JWST Can Do More Science with NIRCam's Short-Wavelength Dispersed Hartmann Sensor

### 2016-08-16 ms.tex
I replaced NIRCam arms with NIRCam channels and a short list of specific recommendations.

\title{Two NIRCam Arms are Better than One: How JWST Can Do More Science with NIRCam's Short-Wavelength Dispersed Hartmann Sensor}

@@

\begin{abstract}
James Webb Space Telescope (JWST) offers unprecedented sensitivity, stability and wavelength coverage for transiting exoplanet studies, opening up new avenues for measuring atmospheric abundances, structure and temperature profiles.
Taking full advantage of JWST spectroscopy of planets from 0.6 $\mu$m to 28$\mu$m, however, will require many observations with a combination of the NIRISS, NIRCam, NIRSpec and MIRI instruments.
In this white paper, we discuss a new NIRCam mode (not yet approved and implemented) that can reduce the number of observations needed for observations in the 1.0$\mu$m to 5.0$\mu$m wavelength range.
The NIRCam instrument was primarily designed as an imager, but includes several grisms for phasing and aligning JWST's 18 hexagonal mirror segments.
NIRCam's long-wavelength arm includes grisms that cover 2.4$\mu$m to 5.0$\mu$m with two observations at $R = 1200 - 1550$ and it has already been approved for wide field slitless spectroscopy.
We propose a new mode that will simultaneously measure spectra for science targets in the 1.0$\mu$m to 2.0$\mu$m range using NIRCam's short-wavelength arm.
This mode, if approved, would take advantage of NIRCam's Dispersed Hartmann Sensor (DHS), which produces 10 spatially separated spectra per source source at $R \sim \DHSresApprox$.
The DHS can observe the brightest host stars of any JWST mode.
We discuss the added benefit of the DHS in constraining abundances in exoplanet atmospheres.
The DHS essentially comes for free with any NIRCam long-wavelength grism observation, but the readout has to be selected to ensure that the long-wavelength grism observations do not saturate and that the high frame rate images on 3 simultaneous detectors does not exceed the data volume JWST can downlink.
Combining both of NIRCam's arms will maximize the science potential of JWST, which is a limited life observatory.
\end{abstract}

@@

No single JWST instrument may measure a spectrum over the full wavelength range from 0.6$\mu$m to 28$\mu$m, so multiple transits or secondary eclipses are necessary to piece together a broadband spectrum.
The largest bandwidths possible are NIRSpec's low resolution prism covering 0.7$\mu$m to 5$\mu$m and MIRI's low resolution prism (LRS) covering 5$\mu$m to 12$\mu$m.
Unfortunately, NIRSpec's low resolution prism mode is so sensitive, it saturates for most sources at a $J$ magnitude of $\sim$11 \citep{beichman2014pasp}, so many nearby bright planet systems must be observed with higher resolution (but smaller wavelength coverage) modes and instruments.
For example, to create a transmission spectrum for HD 209458b \citep{henry00,charbonneau00} from 0.7$\mu$m to 12$\mu$m, an observer would have to stitch together at least four different transits: one with NIRISS (0.7$\mu$m to 2.5$\mu$m), two with NIRCam's long-wavelength arm (2.44$\mu$m to 4.98$\mu$m)\footnote{Or, alternatively NIRSpec's G235H and G395H  grisms covering 1.7~$\mu$ to 5.0$\mu$m.} and one with MIRI (5$\mu$m to 12$\mu$m).
With typical overheads of $\sim$2 hr for 9 hours of pre-ingress, transit and post-ingress, this totals to about ~11 hours per transit or 44 JWST hours.
This was calculated with the Astronomer's Proposal Tools software\footnote{http://www.stsci.edu/hst/proposing/apt} with the NIRISS SOSS mode, NIRCam's F322W2 and F444W grism time series and the MIRI Low Resolution spectrograph (LRS).

The NIRCam instrument is primarily an imager, but it includes several spectroscopic modes originally designed for wavefront sensing.
These include the grisms installed in its long-wavelength arm as well as the Dispersed Hartmann Sensors (DHSs) installed in its short-wavelength arm.
The long-wavelength grisms are used in concert with broadband filters.
The F322W2 filter encompasses 2.44$\mu$m to 4.02$\mu$m half maximum throughput while the F444W filter encompasses 3.88$\mu$m to 4.98$\mu$m at half maximum throughput.
These filters and grism are enabled and supported JWST science modes.
The DHS would be used primarily with the F150W2 filter, which covers from 1.01$\mu$m to 2.35$\mu$m at half maximum throughput.
The range from 2.0$\mu$m to 2.35$\mu$m will probably be discarded because the first order of the spectrum begins to overlap at at 2.0$\mu$m.\footnote{The prism that spatially separate the 10 DHS spectra separate the second order spectrum by 0 to 15 pixels, so the contamination can be reduced in some cases.}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{molecules_DHS_HD209.pdf}
\caption{Model spectra separated by molecule for HD 209458b.
The new DHS mode would allow simultaneous spectra at 1-2$\mu$m ``for free'' when observing with either NIRCam's F322W2 or F444W slitless grism observations in its long-wavelength arm.
The simulated data are for one transit in each of the long-wavelength grism modes, each accompanied by simultaneous DHS grism spectra.
The observations longer than 5$\mu$m are simulated for MIRI's Low Resolution Spectrometer.}\label{fig:209spec}
\end{figure}

The NIRCam instrument can reduce the number of transits needed to build up a transmission or emission spectrum of a planet from 1$\mu$m to 5$\mu$m.
This is because the instrument has a dichroic that splits the incoming beam into a short-wavelength arm and long-wavelength arm, so simultaneous imaging and/or spectroscopy are possible at short ($<$2.4$\mu$m) and long ($>$2.4$\mu$m) wavelengths.
In particular, the DHS can be used to simultaneously cover 1.0$\mu$m to 2.0$\mu$m while the long-wavelength grisms cover either 2.44$\mu$m to 4.02$\mu$m (with the F322W2 filter) or 3.88$\mu$m to 4.98$\mu$m (with the F444W) filter.
Thus, the DHS data comes ``for free'' with any long-wavelength NIRCam grism data, aside from readout and data rate considerations discussed in Section \ref{sec:readout}.
For the example above of HD 209458b, this could save about 11 JWST hours per transit if the wavelengths from 2.0$\mu$m to 2.4$\mu$m and higher Signal to Noise ratios (SNRs) from NIRISS are not required.
The DHS is expected to perform similarly to HST's WFC3 grisms, so it is like a free coordinated HST observation for every long-wavelength NIRCam grism transit.

Since the DHS was not originally designed for science observations, it is not yet an approved mode for the James Webb Space Telescope.
In this white paper, we argue for the inclusion of the DHS mode to fully take advantage of two NIRCam arms to maximize JWST's science potential for transiting planets.
In section \ref{sec:addedScience} we discuss the added science that can be achieved with the DHS, including the brightness limits and improved precision in retrieved parameters for a representative system.
In section \ref{sec:implementation}, we discuss how the DHS would work for time-series observations and the implications of reading out several detectors simultaneously at a high frame rate.

@@

\subsection{DHS properties}
The DHS expands the wavelength range that NIRCam can achieve on a single observation by simultaneously obtaining spectroscopy with NIRCam's short wavelength arm and long wavelength arm.
The maximum wavelength range for the F150W2 filter in the short wavelength arm is fixed at 1.01 to 2.02$\mu$m with the region from 2.02-2.35$\mu$m contaminated by second order overlap.
The DHS are slitless, so the resolution is set by the pupil sizes of the individual DHS elements, which are approximately the same width in the dispersion direction.
Figure \ref{fig:DHSRes} shows the resolution as a function of wavelength, estimated from a narrowly peaked input spectrum and the fits to the spatial width along the spectrum.
There will be about a \DHSgap$\mu$m gap in the spectrum due to the gap between short wavelength detectors.
The center of this gap depends on source position.
The DHS can be used in tandem with either the F322W2 or F444W filter for the long wavelength grism observations.
The properties of the DHS, which can be used simultaneously with the long-wavelength grisms, are listed in Table \ref{tab:DHSgprop}.
If the region from 2-2.23$\mu$m is highly desirable such as for measuring the CH$_4$ feature shown in Figure \ref{fig:HATp12spec}, the DHS can be used with a F200W filter (1.75 to 2.23$\mu$m half maximum) to partially fill in the gap between the short to long wavelength arms.

@@

\begin{figure*}
\centering
\includegraphics[width=0.7\textwidth]{grism_pos_for_max_wcoverage.pdf}
\caption{DHS wavelength coverage for two field positions that will permit grism spectroscopy simultaneously with the F322W2 and F444W filters in the long wavelength arm for the same input source. The V2/V3 axes show the JWST observatory axes and the detectors (A1 - A5) are shown with their (0,0) corners as black boxes. The circles plotted for each grism are the shortest usable wavelength, undispersed wavelength and longest usable wavelength.}\label{fig:DHSwaveWGrism}
\end{figure*}

A point source observed with the DHS and grism modes of NIRCam will produce 10 spectra on the short-wavelength arm and one spectrum on the long-wavelength arm -- see Figure \ref{fig:DHSwaveWGrism}.
The DHS spectra span two of NIRCam's short-wavelength detectors covering $\sim$1 to $\sim$ 2$\mu$m with a \DHSgap$\mu$m gap due to the separation between detectors.
The DHS mode is well suited for one bright source because it spreads the light over the 10 separate spectra, but is ill-suited for a crowded field, where the spectra would overlap.
Each of the DHS spectra are separated in the spatial direction by an average of 123 pixels, so a sub-array of at least 1300 pixels must be extracted to gather all DHS flux for a single source.

@@

%\begin{figure}[!ht]
%\includegraphics[width=0.5\textwidth]{dhs_image_example.png}
%\caption{Example images of the same source observed simultaneously with the DHS on the short-wavelength arm (left) producing 10 spectra and the F444W filter and C grism on the long-wavelength arm (right).
%These images were taken during cryogenic vacuum testing of the NIRCam flight instrument and detectors at NASA Goddard.
%Only 3 detectors will be needed for transit observations \textbf{Update w/ Mod A image and w/ R grism estimate}.}\label{fig:DHSimage}
%\end{figure}

@@

>>>>

\title{Two NIRCam channels are Better than One: How JWST Can Do More Science with NIRCam's Short-Wavelength Dispersed Hartmann Sensor}

@@

\begin{abstract}
James Webb Space Telescope (JWST) offers unprecedented sensitivity, stability and wavelength coverage for transiting exoplanet studies, opening up new avenues for measuring atmospheric abundances, structure and temperature profiles.
Taking full advantage of JWST spectroscopy of planets from 0.6 $\mu$m to 28$\mu$m, however, will require many observations with a combination of the NIRISS, NIRCam, NIRSpec and MIRI instruments.
In this white paper, we discuss a new NIRCam mode (not yet approved and implemented) that can reduce the number of observations needed for observations in the 1.0$\mu$m to 5.0$\mu$m wavelength range.
The NIRCam instrument was primarily designed as an imager, but includes several grisms for phasing and aligning JWST's 18 hexagonal mirror segments.
NIRCam's long-wavelength channel includes grisms that cover 2.4$\mu$m to 5.0$\mu$m with two observations at $R = 1200 - 1550$ and it has already been approved for wide field slitless spectroscopy.
We propose a new mode that will simultaneously measure spectra for science targets in the 1.0$\mu$m to 2.0$\mu$m range using NIRCam's short-wavelength channel.
This mode, if approved, would take advantage of NIRCam's Dispersed Hartmann Sensor (DHS), which produces 10 spatially separated spectra per source source at $R \sim \DHSresApprox$.
The DHS can observe the brightest host stars of any JWST mode.
We discuss the added benefit of the DHS in constraining abundances in exoplanet atmospheres.
The DHS essentially comes for free with any NIRCam long-wavelength grism observation, but the readout has to be selected to ensure that the long-wavelength grism observations do not saturate and that the high frame rate images on 3 simultaneous detectors does not exceed the data volume JWST can downlink.
Combining both of NIRCam's channels will maximize the science potential of JWST, which is a limited life observatory.
\end{abstract}

@@

No single JWST instrument may measure a spectrum over the full wavelength range from 0.6$\mu$m to 28$\mu$m, so multiple transits or secondary eclipses are necessary to piece together a broadband spectrum.
The largest bandwidths possible are NIRSpec's low resolution prism covering 0.7$\mu$m to 5$\mu$m and MIRI's low resolution prism (LRS) covering 5$\mu$m to 12$\mu$m.
Unfortunately, NIRSpec's low resolution prism mode is so sensitive, it saturates for most sources at a $J$ magnitude of $\sim$11 \citep{beichman2014pasp}, so many nearby bright planet systems must be observed with higher resolution (but smaller wavelength coverage) modes and instruments.
For example, to create a transmission spectrum for HD 209458b \citep{henry00,charbonneau00} from 0.7$\mu$m to 12$\mu$m, an observer would have to stitch together at least four different transits: one with NIRISS (0.7$\mu$m to 2.5$\mu$m), two with NIRCam's long-wavelength channel (2.44$\mu$m to 4.98$\mu$m)\footnote{Or, alternatively NIRSpec's G235H and G395H  grisms covering 1.7~$\mu$ to 5.0$\mu$m.} and one with MIRI (5$\mu$m to 12$\mu$m).
With typical overheads of $\sim$2 hr for 9 hours of pre-ingress, transit and post-ingress, this totals to about ~11 hours per transit or 44 JWST hours.
This was calculated with the Astronomer's Proposal Tools software\footnote{http://www.stsci.edu/hst/proposing/apt} with the NIRISS SOSS mode, NIRCam's F322W2 and F444W grism time series and the MIRI Low Resolution spectrograph (LRS).

The NIRCam instrument is primarily an imager, but it includes several spectroscopic modes originally designed for wavefront sensing.
These include the grisms installed in its long-wavelength channel as well as the Dispersed Hartmann Sensors (DHSs) installed in its short-wavelength channel.
The long-wavelength grisms are used in concert with broadband filters.
The F322W2 filter encompasses 2.44$\mu$m to 4.02$\mu$m half maximum throughput while the F444W filter encompasses 3.88$\mu$m to 4.98$\mu$m at half maximum throughput.
These filters and grism are enabled and supported JWST science modes.
The DHS would be used primarily with the F150W2 filter, which covers from 1.01$\mu$m to 2.35$\mu$m at half maximum throughput.
The range from 2.0$\mu$m to 2.35$\mu$m will probably be discarded because the first order of the spectrum begins to overlap at at 2.0$\mu$m.\footnote{The prism that spatially separate the 10 DHS spectra separate the second order spectrum by 0 to 15 pixels, so the contamination can be reduced in some cases.}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{molecules_DHS_HD209.pdf}
\caption{Model spectra separated by molecule for HD 209458b.
The new DHS mode would allow simultaneous spectra at 1-2$\mu$m ``for free'' when observing with either NIRCam's F322W2 or F444W slitless grism observations in its long-wavelength channel.
The simulated data are for one transit in each of the long-wavelength grism modes, each accompanied by simultaneous DHS grism spectra.
The observations longer than 5$\mu$m are simulated for MIRI's Low Resolution Spectrometer.}\label{fig:209spec}
\end{figure}

The NIRCam instrument can reduce the number of transits needed to build up a transmission or emission spectrum of a planet from 1$\mu$m to 5$\mu$m.
This is because the instrument has a dichroic that splits the incoming beam into a short-wavelength channel and long-wavelength channel, so simultaneous imaging and/or spectroscopy are possible at short ($<$2.4$\mu$m) and long ($>$2.4$\mu$m) wavelengths.
In particular, the DHS can be used to simultaneously cover 1.0$\mu$m to 2.0$\mu$m while the long-wavelength grisms cover either 2.44$\mu$m to 4.02$\mu$m (with the F322W2 filter) or 3.88$\mu$m to 4.98$\mu$m (with the F444W) filter.
Thus, the DHS data comes ``for free'' with any long-wavelength NIRCam grism data, aside from readout and data rate considerations discussed in Section \ref{sec:readout}.
For the example above of HD 209458b, this could save about 11 JWST hours per transit if the wavelengths from 2.0$\mu$m to 2.4$\mu$m and higher Signal to Noise ratios (SNRs) from NIRISS are not required.
The DHS is expected to perform similarly to HST's WFC3 grisms, so it is like a free coordinated HST observation for every long-wavelength NIRCam grism transit.

Since the DHS was not originally designed for science observations, it is not yet an approved mode for the James Webb Space Telescope.
In this white paper, we argue for the inclusion of the DHS mode to fully take advantage of two NIRCam channels to maximize JWST's science potential for transiting planets.
In section \ref{sec:addedScience} we discuss the added science that can be achieved with the DHS, including the brightness limits and improved precision in retrieved parameters for a representative system.
In section \ref{sec:implementation}, we discuss how the DHS would work for time-series observations and the implications of reading out several detectors simultaneously at a high frame rate.

@@

\subsection{DHS properties}
The DHS expands the wavelength range that NIRCam can achieve on a single observation by simultaneously obtaining spectroscopy with NIRCam's short wavelength channel and long wavelength channel.
The maximum wavelength range for the F150W2 filter in the short wavelength channel is fixed at 1.01 to 2.02$\mu$m with the region from 2.02-2.35$\mu$m contaminated by second order overlap.
The DHS are slitless, so the resolution is set by the pupil sizes of the individual DHS elements, which are approximately the same width in the dispersion direction.
Figure \ref{fig:DHSRes} shows the resolution as a function of wavelength, estimated from a narrowly peaked input spectrum and the fits to the spatial width along the spectrum.
There will be about a \DHSgap$\mu$m gap in the spectrum due to the gap between short wavelength detectors.
The center of this gap depends on source position.
The DHS can be used in tandem with either the F322W2 or F444W filter for the long wavelength grism observations.
The properties of the DHS, which can be used simultaneously with the long-wavelength grisms, are listed in Table \ref{tab:DHSgprop}.
If the region from 2-2.23$\mu$m is highly desirable such as for measuring the CH$_4$ feature shown in Figure \ref{fig:HATp12spec}, the DHS can be used with a F200W filter (1.75 to 2.23$\mu$m half maximum) to partially fill in the gap between the short to long wavelength channels.

@@

\begin{figure*}
\centering
\includegraphics[width=0.7\textwidth]{grism_pos_for_max_wcoverage.pdf}
\caption{DHS wavelength coverage for two field positions that will permit grism spectroscopy simultaneously with the F322W2 and F444W filters in the long wavelength channel for the same input source. The V2/V3 axes show the JWST observatory axes and the detectors (A1 - A5) are shown with their (0,0) corners as black boxes. The circles plotted for each grism are the shortest usable wavelength, undispersed wavelength and longest usable wavelength.}\label{fig:DHSwaveWGrism}
\end{figure*}

A point source observed with the DHS and grism modes of NIRCam will produce 10 spectra on the short-wavelength channel and one spectrum on the long-wavelength channel -- see Figure \ref{fig:DHSwaveWGrism}.
The DHS spectra span two of NIRCam's short-wavelength detectors covering $\sim$1 to $\sim$ 2$\mu$m with a \DHSgap$\mu$m gap due to the separation between detectors.
The DHS mode is well suited for one bright source because it spreads the light over the 10 separate spectra, but is ill-suited for a crowded field, where the spectra would overlap.
Each of the DHS spectra are separated in the spatial direction by an average of 123 pixels, so a sub-array of at least 1300 pixels must be extracted to gather all DHS flux for a single source.

@@

%\begin{figure}[!ht]
%\includegraphics[width=0.5\textwidth]{dhs_image_example.png}
%\caption{Example images of the same source observed simultaneously with the DHS on the short-wavelength channel (left) producing 10 spectra and the F444W filter and C grism on the long-wavelength channel (right).
%These images were taken during cryogenic vacuum testing of the NIRCam flight instrument and detectors at NASA Goddard.
%Only 3 detectors will be needed for transit observations \textbf{Update w/ Mod A image and w/ R grism estimate}.}\label{fig:DHSimage}
%\end{figure}

@@

\subsubsection{Specific Changes Needed for Implementation}
We summarize the specific changes to the existing operations needed for the DHS here:
\begin{enumerate}
\item Allow the NIRCam pupil wheel to be used for science at the DHS element position with spectra parallel with detector rows
\item Allow subarray locations to be different for the short wavelength arrays and the long wavelength arrays in order to capture the spectrum for Element numbers 7 and 8 for subarrays
\item Choose field points from NIRCam's long wavelength grism to allow simultaneous DHS spectroscopy with all 10 DHS elements
\item Change the OSS software and timing so that large subarrays may used with the DHS simultaneously with small 2048$\times$64 subarrays with the long wavelength grisms
\end{enumerate}
